---
title: "2016 Presidential Donations"
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: fill
    theme: flatly
runtime: shiny
---

```{r setup, include=FALSE}
# Packages
library(flexdashboard)
library(shiny)
library(tidyverse)
library(plotly)
library(scales)
library(forcats)
library(DT)
library(lubridate)
library(leaflet)
library(maps)

# GGPLOT Theme
theme_set(theme_bw())

# Data Import
df <- read_csv("/Users/bi539pd/Google Drive/Work/Personal Projects/2016 Presidential Campaign/App/forapp.csv")
#df <- read_csv("P00000001-ALL.csv")

# Data Clean- get rid of negative donations
df$contb_receipt_amt[df$contb_receipt_amt <0] <- NA
df <- df[!is.na(df$contb_receipt_amt),]
```

Donation Statistics {data-orientation=rows}
=====================================  

Sidebar {.sidebar}
------------------------------------
**Introduction**

This application uses data provided from the Federal Election Committee from the 2016 Presidential election. Donations to candidates are recorded with the contributer's name, city, state, zip, employer, occupation, amount, and date of receipt.

This page lets you compare donation statistics.

```{r}
# Rennder selecter for page 1
renderUI({
  selectizeInput("cand_select",
                 "Select Candidates",
                 choices=unique(df$cand_nm),
                 multiple=T,
                 selected=NULL)
})

# Reactive dataset for page 1
df1 <- reactive({
  filter(df, cand_nm %in% input$cand_select)
})
```

Row
---------------------------

### Total Donations (Millions of Dollars)
```{r}
renderPlotly({
  validate(
    need(length(input$cand_select)>0,"Select Candidates")
  )
  df1 <- df1() %>%
    group_by(cand_nm) %>%
    summarize("Total Donations" = sum(contb_receipt_amt)/1000000) 
  
    df1$cand_nm <- fct_reorder(df1$cand_nm,df1$`Total Donations`)
    
    ggplot(df1, aes(x=cand_nm, y=`Total Donations`))+
    geom_bar(fill="blue",stat="identity")+
    coord_flip()+
    scale_y_continuous(label=dollar)+
    labs(x=NULL,y=NULL)
})

```


### Number of unique donors (Thousands)
```{r}
renderPlotly({
  validate(
    need(length(input$cand_select)>0,"Select Candidates")
  )
  df1 <- df1() %>%
    group_by(cand_nm) %>%
    summarise("Donors" = length(unique(contbr_nm))/1000)
  
  df1$cand_nm <- fct_reorder(df1$cand_nm,df1$Donors)
  
   ggplot(df1, aes(x=cand_nm, y=Donors))+
    geom_bar(fill="blue",stat="identity")+
    coord_flip()+
    labs(x=NULL,y=NULL)

})
```


Row {.tabset .tabset-fade}
--------------------------
### Donor Statistics
```{r}
totals <- reactive({ 
  totals <- df1() %>%
  group_by(cand_nm) %>%
  summarise("Total (Millions)" = sum(contb_receipt_amt)/1000000,
            "Number of Donations" = n(),
            "Highest Donation" = max(contb_receipt_amt),
            "Median Donation" = median(contb_receipt_amt),
            "Average Donation" = mean(contb_receipt_amt))

totals[,2:6] <- apply(totals[,2:6],2,
                      function(x) format(x, big.mark=",",
                                         digits=2))
totals
})

renderDataTable({
    validate(
    need(length(input$cand_select)>0,"Select Candidates")
  )
  datatable(totals(),options=list(searching=FALSE,paging=FALSE))
})


```

### Donations Over Time
```{r}
# Totals by Month
df$contb_receipt_dt <- dmy(df$contb_receipt_dt)
df$Month <- floor_date(df$contb_receipt_dt,"month")

df.m <- reactive({
df.m <-  df1() %>%
  group_by(cand_nm,Month) %>%
  summarize(Total = sum(contb_receipt_amt)) %>%
  mutate(cumsum = cumsum(Total))

df.m$cumsum <- round(df.m$cumsum/1000000,2)
df.m$cumsum[df.m$cumsum==0] <- NA
df.m <- na.omit(df.m)
df.m
})

# Plot
renderPlotly({
    validate(
    need(length(input$cand_select)>0,"Select Candidates")
  )
ggplot(df.m(), aes(x=Month,y=cumsum,col=cand_nm))+
  geom_line() +
  labs(x=NULL,y="Total Donations (Millions)",
       col="candidate")
  })

```

Donors: Individuals {data-orientation=rows}
=====================================  

Sidebar {.sidebar}
------------------------------------
**Introduction**

This application uses data provided from the Federal Election Committee from the 2016 Presidential election. Donations to candidates are recorded with the contributer's name, city, state, zip, employer, occupation, amount, and date of receipt.

This page lets you look at the top 500 individual donors.

```{r}
donors <- df %>%
  group_by(contbr_nm) %>%
  summarise(Total= sum(contb_receipt_amt)) %>%
  arrange(desc(Total))

donors <- donors$contbr_nm[1:500]

renderUI({
  selectizeInput("donors_select",
                 "Select Donors",
                 choices=donors,
                 multiple=TRUE,
                 selected=NULL)
})

df.d <- reactive({
  filter(df, contbr_nm %in% input$donors_select)
})

```

Row
-------------------------------
### Sum Donated
```{r}
renderValueBox({
  valueBox("Total",
           paste("$",format(sum(df.d()$contb_receipt_amt),big.mark=","),
                 sep=""),
           icon="fa-money",
           color="lightgreen")
})
```

### Number Donations
```{r}
renderValueBox({
  valueBox("# Donations",
           nrow(df.d()),
           icon="fa-plus-circle",
           color="lightblue")
})
```

### Different Candidates
```{r}
renderValueBox({
  valueBox("# Candidates",
           length(unique(df.d()$cand_nm)),
           icon="fa-group",
           color="lightblue")
})

```

Row {.tabset .tabset-fade}
--------------------
### Donor Statistics
```{r}
renderDataTable({
    validate(
    need(length(input$donors_select)>0, "Select Donor(s)")
  )
  df.d <- df.d() %>%
    group_by(contbr_nm) %>%
    summarise("Number of Donations" = n(),
              "Min Donation" = min(contb_receipt_amt),
              "Mean Donation" = mean(contb_receipt_amt),
              "Max Donation" = max(contb_receipt_amt))
  df.d$`Mean Donation` <- format(df.d$`Mean Donation`,
                                 digits=2,
                                 big.mark=",")
  df.d$`Max Donation` <- format(df.d$`Max Donation`,
                                 digits=2,
                                 big.mark=",")
  df.d$`Min Donation` <- format(df.d$`Min Donation`,
                                 digits=2,
                                 big.mark=",")

  datatable(df.d,options=list(searching=FALSE))
  
})

```

### Donor Trends
```{r}
renderPlotly({
      validate(
    need(length(input$donors_select)>0, "Select Donor(s)")
  )
df.d() %>%
  group_by(contbr_nm, cand_nm, Month) %>%
  summarise(Sum= sum(contb_receipt_amt)) %>%
  ggplot(., aes(x=Month,y=Sum,col=contbr_nm)) +
  geom_point() +
  geom_line() +
  facet_wrap(~cand_nm)+
    labs(y="Donation Amount",
         x=NULL,
         col="Donor") +
  theme(axis.text.x=element_text(angle=90, hjust=1))
})
  
```


Donors: Employers {data-orientation=rows}
=====================================  

Sidebar {.sidebar}
------------------------------------
**Introduction**

This application uses data provided from the Federal Election Committee from the 2016 Presidential election. Donations to candidates are recorded with the contributer's name, city, state, zip, employer, occupation, amount, and date of receipt.

This page lets you look at the top 500 employers represented by donors.

```{r}
donors.e <- df %>%
  group_by(contbr_employer) %>%
  summarise(Total= sum(contb_receipt_amt)) %>%
  arrange(desc(Total))

donors.e <- donors.e$contbr_employer[1:500]

renderUI({
  selectizeInput("employers_select",
                 "Select Employers",
                 choices=donors.e,
                 multiple=TRUE,
                 selected=NULL)
})

df.e <- reactive({
  filter(df, contbr_employer %in% input$employers_select)
})

```

Row
-------------------------------
### Sum Donated
```{r}
renderValueBox({
  valueBox("Total",
           paste("$",format(sum(df.e()$contb_receipt_amt),big.mark=","),
                 sep=""),
           icon="fa-money",
           color="lightgreen")
})
```

### Number Donations
```{r}
renderValueBox({
  valueBox("# Donations",
           nrow(df.e()),
           icon="fa-plus-circle",
           color="lightblue")
})
```

### Different Candidates
```{r}
renderValueBox({
  valueBox("# Candidates",
           length(unique(df.e()$cand_nm)),
           icon="fa-group",
           color="lightblue")
})

```

Row {.tabset .tabset-fade}
--------------------
### Donor Statistics
```{r}
renderDataTable({
    validate(
    need(length(input$donors_select)>0, "Select Donor(s)")
  )
  df.e <- df.e() %>%
    group_by(contbr_employer) %>%
    summarise("Number of Donations" = n(),
              "Min Donation" = min(contb_receipt_amt),
              "Mean Donation" = mean(contb_receipt_amt),
              "Max Donation" = max(contb_receipt_amt))
  df.e$`Mean Donation` <- format(df.e$`Mean Donation`,
                                 digits=2,
                                 big.mark=",")
  df.e$`Max Donation` <- format(df.e$`Max Donation`,
                                 digits=2,
                                 big.mark=",")
  df.e$`Min Donation` <- format(df.e$`Min Donation`,
                                 digits=2,
                                 big.mark=",")

  datatable(df.e,options=list(searching=FALSE))
  
})

```

### Donor Trends
```{r}
renderPlotly({
      validate(
    need(length(input$employers_select)>0, "Select Donor(s)")
  )
df.e() %>%
  group_by(contbr_employer, cand_nm, Month) %>%
  summarise(Sum= sum(contb_receipt_amt)) %>%
  ggplot(., aes(x=Month,y=Sum,col=contbr_employer)) +
  geom_point() +
  geom_line() +
  facet_wrap(~cand_nm)+
    labs(y="Donation Amount",
         x=NULL,
         col="Employer") +
  theme(axis.text.x=element_text(angle=90, hjust=1))
})
  
```

Geography {data-orientation=rows}
=====================================  

Sidebar {.sidebar}
------------------------------------
**Introduction**

This application uses data provided from the Federal Election Committee from the 2016 Presidential election. Donations to candidates are recorded with the contributer's name, city, state, zip, employer, occupation, amount, and date of receipt.

This page lets you compare donations geographically

```{r}
states <- df
```

Row
---------------------------
### Map
```{r}
renderLeaflet({
leaflet() %>%
  addTiles() %>%  # Add default OpenStreetMap map tiles
  addMarkers(lng=174.768, lat=-36.852, popup="The birthplace of R")
  })
```





